What the hell is the title of this paper?

### 2018-05-14 NIPS 2018/main.tex
Small edits

Recent advancements have allowed reinforcement learning (RL) \cite{sutton1998reinforcement} to achieve impressive results in a wide variety of complex tasks, ranging from Atari \cite{mnih2015human}, to the game of Go \cite{silver2016mastering}, to the control of sophisticated robotics systems \cite{kober2009policy,lillicrap2015continuous,levine2016end}. The main limitation is that RL algorithms still require an enormous amount of experience samples before successfully learning such complicated tasks. One of the most promising solutions is transfer learning, which focuses on reusing past knowledge available to the agent in order to reduce the sample-complexity for learning new tasks. In the typical settings of transfer in RL \cite{taylor2009transfer}, the agent is assumed to have already solved a set of \textit{source tasks} generated from some unknown distribution. Then, given a \textit{target task} drawn from the same distribution, or a slightly different one, the agent can rely on knowledge from the source tasks to speed up the learning process. This constitutes a significant advantage over plain RL, where the agent learns each new task from scratch independently of previous learning experience. Several algorithms have been proposed in the literature to transfer experience samples \cite{lazaric2008transfer,taylor2008transferring}, policies/options \cite{fernandez2006probabilistic,konidaris2007building}, rewards \cite{konidaris2006autonomous}, features \cite{barreto2017successor}, parameters \cite{doshi2016hidden,killian2017robust}, and so on. We refer the reader to \cite{taylor2009transfer} for a thorough survey on transfer in RL.

@@

Under the assumption that tasks follow a certain distribution, an intuitive choice for designing a transfer algorithm is to attempt at characterizing the uncertainty over the target task. Then, an ideal algorithm would leverage prior knowledge from the source tasks to interact with the target task in such a way that this uncertainty is reduced as quickly as possible. This simple intuition makes Bayesian \todo{probabilistic? distributional?} methods appealing approaches for transfer in RL, and many previous works have been proposed in this direction. \cite{wilson2007multi} assume tasks share similarities in their dynamics and rewards and propose a hierarchical Bayesian model for the distribution of these two elements. Similarly, \cite{lazaric2010bayesian} assume tasks are similar in their value functions and design a different hierarchical Bayesian model for transferring such information. More recently, \cite{doshi2016hidden}, and its extension \cite{killian2017robust}, consider tasks whose dynamics are governed by some hidden parameters, and propose efficient Bayesian models for quickly learning such parameters in new tasks. However, all these algorithms require specific, and sometimes restrictive, assumptions (e.g., on the distributions involved or the function approximators adopted), which might limit their practical applicability. The importance of having transfer algorithms that alleviate the need of strong assumptions and that can be easily adapted to different contexts motivates us to take a more general approach.

Similarly to \cite{lazaric2010bayesian}, we assume tasks to share similarities in their value functions and use the given source tasks to learn a distribution over such functions. Then, we use this distribution as a prior for learning the target task and we propose and efficient variational approximation of the corresponding posterior. Leveraging on recent ideas from randomized value functions \cite{osband2014generalization}, we design a Thompson Sampling-based algorithm which efficiently explores the target task by repeatedly sampling from the posterior and acting greedily w.r.t. (with respect to) the sampled value function. We show that our approach is very general, in the sense that it does not require any specific choice of function approximator or prior/posterior distribution models. We propose a finite-sample analysis that theoretically validates our approach, while empirically evaluating its performance on three domains with increasing level of difficulty.

@@

We consider a distribution $\mathcal{D}$ over tasks, where each task is modeled as a discounted Markov Decision Process (MDP). We define an MDP as a tuple $\mathcal{M} = \langle \mathcal{S},\mathcal{A},\mathcal{P},\mathcal{R},p_0,\gamma\rangle$\todo{Should we provide a more transfer-oriented definition?}, where $\mathcal{S}$ is the state-space, $\mathcal{A}$ is a finite set of actions, $\mathcal{P}(\cdot | s,a)$ is the distribution of the next state $s'$ given that action $a$ is taken in state $s$, $\mathcal{R}: \mathcal{S}\times\mathcal{A}\rightarrow\mathbb{R}$ is the reward function, $p_0$ is the initial-state distribution, and $\gamma\in [0,1)$ is the discount factor. We assume the reward function to be uniformly bounded by a constant $R_{max}>0$. A deterministic policy $\pi : \mathcal{S} \rightarrow \mathcal{A}$ is a mapping from states to actions. At the beginning of each episode of interaction, the initial state $s_0$ is drawn from $p_0$. Then, the agent takes the action $a_0 = \pi(s_0)$, receives a reward $\mathcal{R}(s_0,a_0)$, transitions to the next state $s_1 \sim \mathcal{P}(\cdot | s_0,a_0)$, and the process is repeated. The goal is to find the policy maximizing the long-term return over a possibly infinite horizon: $\max_{\pi}J(\pi)\triangleq\mathbb{E}[\sum_{t=0}^{\infty} \gamma^t r_t \mid \mathcal{M},\pi]$. To this end, we define the optimal value function $Q^*(s,a)$ as the expected return obtained by taking action $a$ in state $s$ and following an optimal policy thereafter. Then, an optimal policy $\pi^*$ is a policy that is greedy with respect to the optimal value function, i.e., $\pi^*(s) = \argmax_a Q^*(s,a)$ for all states $s$. It can be shown (e.g., \cite{puterman1994markov}) that $Q^*$ is the unique fixed-point of the optimal Bellman operator $T$ defined by $TQ(s,a) = \mathcal{R}(s,a) + \gamma\mathbb{E}_{\mathcal{P}}[\max_{a'}Q(s',a')]$ for any value function $Q$. From now on, we adopt the term $Q$-function to denote any plausible value function, i.e., any function $Q : \mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ uniformly bounded by $\frac{R_{max}}{1-\gamma}$.

@@

Our approach is mostly related to \cite{lazaric2010bayesian}. Although we both assume the tasks to share similarities in their value functions, \cite{lazaric2010bayesian} consider only linear approximators and adopt a hierarchical Bayesian model of the corresponding weights' distribution, which is assumed Gaussian. On the other hand, our variational approximation allows for more general distribution families and can be combined with non-linear approximators. Furthermore, \cite{lazaric2010bayesian} propose a Dirichlet process model for the case where weights cluster into different classes, which relates to our mixture formulation and proves again the importance of capturing more complicated task distributions. In \cite{wilson2007multi}, the authors propose a hierarchical Bayesian model for the distribution over tasks. Differently from our approach and \cite{lazaric2010bayesian}, they consider a distribution over transition probabilities and rewards, rather than value functions. In the same spirit of our method, they consider a Thompson sampling-based procedure which, at each iteration, samples a new task from the posterior and solves it. However, \cite{wilson2007multi} consider only finite MDPs, which poses a severe limitation on the algorithm's applicability. On the contrary, our approach can handle high-dimensional tasks. In \cite{doshi2016hidden}, the authors consider a family of tasks whose dynamics are governed by some hidden parameters and use Gaussian processes (GPs) to model such dynamics across tasks. This approach was recently extended in \cite{killian2017robust}, where the authors replace GPs with Bayesian neural networks to obtain a more scalable approach.\todo{More comments about HiP-MDPs? Should we discuss more related works?}

In the RL community, our approach is related to value function randomization\cite{osband2014generalization}. The authors extend the well-known LSTD \cite{boyan1999least} by adopting Bayesian linear regression to model the uncertainty over the predicted value function weights, and use that to perform a form of Thompson sampling. Such approach was recently extended by \cite{azizzadenesheli2018efficient}, where the approximator is replaced by a Bayesian neural network, leading to an algorithm capable of solving much more complicated problems. Both these algorithms rely on the Gaussian assumption and, since they work in plain RL settings, have no informative prior available. On the other hand, our variational approximation allows more complex distributions (e.g., mixtures) to be adopted, while knowledge from the source tasks allows us to learn very informative priors.

@@

In this section, we provide an experimental evaluation of our approach. We begin by analyzing the behavior of our algorithm on a toy problem in Sec. \ref{sec:gw}. Then, in Sec. \ref{sec:cc}, we consider two classic control benchmarks, namely Cartpole and Mountain Car. We conclude with a complex maze navigation task in Sec. \ref{sec:mn}. In all experiments, we compare our Gaussian variational transfer algorithm (GVT) and the version using mixture of Gaussians with $c$ components ($c$-MGVT) to plain no-transfer RL (NT) with $\epsilon$-greedy exploration. To the best of our knowledge, no existing transfer algorithm is directly comparable to our approach from an experimental perspective. \todo{Any better motivation?}

@@

\begin{figure*}[t] 
  \begin{subfigure}[b]{0.15\textwidth}
    \includegraphics[trim=0.0cm 0.2cm 0.0cm 0.2cm,clip=true,height=3cm, width=2.0cm]{images/rooms2.pdf}
    \caption{}
  \end{subfigure}
  \begin{subfigure}[b]{0.27\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=2.9cm]{images/three-room/lrev.pdf}
    \caption{}
  \end{subfigure}
    \begin{subfigure}[b]{0.27\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.6cm 1.3cm,clip=true,height=2.9cm]{images/three-room-gen/lrev.pdf}
    \caption{}
  \end{subfigure}
    \begin{subfigure}[b]{0.27\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.6cm 1.3cm,clip=true,height=2.9cm]{images/likelihood/lrew.pdf}
    \caption{}
  \end{subfigure}
  \caption{(a) the rooms environment, (b) transfer from $10$ source tasks with both doors moving, (c) transfer from $10$ source tasks with only one door moving, and (d) transfer performance as a function of how likely the target task is according to the prior.}
  \label{fig:dipfdbsgraphplots}
\end{figure*}

@@

In order to prove that our guesses are correct, we generate a set of $50$ source tasks for the three-room environment of Fig. \ref{} by sampling both door positions uniformly, and solve all of them by directly minimizing the TD error as presented in Sec. \ref{sec:td}. In order to make sure that their solutions are accurate enough, we allow sampling the initial state uniformly in the environment and run until convergence. Then, we use our algorithms to transfer from $10$ source tasks sampled from the previously generated set. The average return over the last $50$ learning episodes as a function of the number of iterations is shown in Fig. \ref{}. Each curve is the result of $20$ independent runs, each resampling the target and source tasks. $95\%$ confidence intervals are shown. Further details on the parameters adopted in this experiment are given in App. \ref{}. As expected, the no-transfer (NT) algorithm fails at learning the task in so few iterations due to the limited exploration provided by an $\epsilon$-greedy policy. On the other hand, all our algorithms achieve a significant speed-up and are able to converge to the optimal performance in few iterations, with GVT being slightly slower. Interestingly, we notice that there is no advantage in adopting more than $1$ component for the posterior in MGVT. This is intuitive since, when the algorithm quickly figures out which task is being solved, it will move all components in the same direction.

To better understand the differences between GVT and MGVT, we now consider transferring from a slightly different distribution than the one from which target tasks are drawn. We generate again $50$ source tasks but this time with the bottom door fixed in the center and the other one moving. Then, we repeat the previous experiment, allowing both doors to move when sampling target tasks. The results are shown in Fig. \ref{}. Interestingly, MGVT seems almost unaffected by this change, proving that is has sufficient representation power to generalize to slightly different task distributions. The same does not hold for GVT, which now is not able to solve many of the sampled target tasks,thus the very high variance. This proves again that assuming Gaussian distributions can pose severe limitations in our transfer settings.

Finally, we analyze the transfer performance as a function of how likely the target task is according to the prior. We consider a two-room version of the environment of Fig. \ref{}. Differently from before, we generate tasks by sampling the door position from a Gaussian with mean $5$, and standard deviation $1.8$, so that tasks where the door is near the sides are very unlikely. Fig. \ref{} shows the performance reached by GVT and MGVT with $1$ component at fixed iterations as a function of how likely the target task is according to such distribution. As expected GVT achieves poor performance on very unlikely tasks, even after many iterations. In fact, estimating a single Gaussian distribution definitely entails some information loss, especially about the unlikely tasks. On the other hand, MGVT keeps such information and, consequently, performs much better. Perhaps not surprisingly, MGVT reaches the optimal performance in $4k$ iterations no matter what task is being solved.

@@

\begin{figure*}[t]
  \begin{subfigure}[b]{0.15\textwidth}
    \includegraphics[trim=0.0cm 0.0cm 0.0cm 0.0cm,clip=true,height=3cm, width=2.0cm]{images/maze/maze13.png}
    \caption{}
  \end{subfigure}
  \begin{subfigure}[b]{0.27\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=2.9cm]{images/cartpole/lrev.pdf}
    \caption{}
  \end{subfigure}
    \begin{subfigure}[b]{0.27\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.6cm 1.3cm,clip=true,height=2.9cm]{images/mountaincar/lrev.pdf}
    \caption{}
  \end{subfigure}
    \begin{subfigure}[b]{0.27\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.6cm 1.3cm,clip=true,height=2.9cm]{images/likelihood/lrew.pdf}
    \caption{}
  \end{subfigure}
  \caption{(a) Sample Maze, (b) transfer from $10$ sources in Cartpole, (c) transfer from $10$ source tasks in Mountain Car, and (d) transfer from $10$ sources in Mazes 0, 6, 13---see App. \ref{}}
  \label{fig:ccmazeplots}
\end{figure*}

To evaluate the performance of our transfer algorithms in more common benchmarks, we chose to use the well-known classic control environments: Cartpole and Mountain Car as defined in \citep{sutton1998reinforcement}. In the case of Cartpole we generated $20$ source tasks by uniformly sampling the cart mass, the pole mass and length. Instead, for Mountain Car, we generated the $20$ sources by uniformly sampling the base speed of the car. We kept a discount factor of $\gamma=0.99$ for both environments. Moreover, to train the source tasks we chose to use a Multilayer Perceptron (MLP) as approximator. We trained all source tasks until convergence and to transfer, in both settings, we used GVT and 1-MGVT choosing randomly $10$ of the trained sources.

In Fig. \ref{}, we show the results of transfer in Cartpole from which it is easily noticed a near-optimal jump-start in both GVT and 1-MGVT that during the following iterations converges to the optimal performance. In the other hand, transfer in Mountain Car as shown in Fig. \ref{}, a more complicated task than Cartpole, causes GVT to struggle to solve the target task but 1-MGVT reaches convergence within the same number of iterations to the optimal performances.

@@

For this setting we propose a maze environment, such as those in Fig. \ref{}, for a robotic agent to navigate. We created a set of 20 mazes of size $10\times10$ with continuous space within which we ensured four groups: each group with the goal position in one of the corners of the maze. In order to simulate a robotic agent, we set an action space that allows to control the linear and angular motion by a fixed distance of $0.5$ (\emph{move forward}) and a fixed angle of $\pm 22.5º$ (\emph{rotate counter- and clockwise}). Moreover, we enhanced the state space to include, in addition to the absolute position and orientation, a set of distance measures to the nearest obstacles from the agent's perspective with maximum range of $2$ and a field of vision from $-90º$ to $90º$ relative to its orientation. In this way, we provide the agent with information that a robot would normally collect from LIDARs, sonars and other sensors alike. Finally, in the same field of perception, we provide a boolean vector indicating whether the goal is within the range of observation. Regarding rewards, the agent gets 1 whenever it enters the goal area and 0 in any other situation and the MDP has a discount factor of $\gamma = 0.99$.

The main motivation for this set-up is that a robot is able to exploit locally-sensed information in addition to its estimation of the actual position---as opposed to the Rooms problem presented above---and it is such information that the robotic agent could learn to use and, also, exploit in different maze configurations leveraging its previous experience. Moreover, simple variations in the maze structure could, potentially, produce big differences in the value function with respect to the ones seen in previous tasks which makes the setting very good to assess how robust our algorithm is to negative transfer.

In App \ref the mazes used are enumerated. These were solved using Deep Q Networks (DQN) [] \todo{cite DQN?} with a multilayer perceptron with hidden layers $32\times32$  and, in order to perform the knowledge transfer, the target maze would be excluded from the source tasks given to the algorithm. Finally, the MGVT algorithm, with one component for the posterior distribution, was used to solve the target tasks shown in Fig. *.

@@

Under Gaussian distributions, all quantities of interest for using Alg. \ref{alg:transfer} can be computed very easily. The KL divergence between the prior and approximate posterior can be computed in closed-form as:
\begin{equation}
KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \frac{1}{2}\left( \log\frac{\left\lvert \bm{\Sigma}_p \right\rvert}{\left\lvert \bm{\Sigma} \right\rvert} + \text{Tr}\left( \bm{\Sigma}_p^{-1}\bm{\Sigma} \right) + (\bm{\mu} - \bm{\mu}_p)^T \bm{\Sigma}_p^{-1} (\bm{\mu} - \bm{\mu}_p) -K \right)
\end{equation}
for $\bm{\xi} = (\bm{\mu},\bm{L})$ and $\bm{\Sigma} = \bm{LL}^T$. Its gradients with respect to the variational parameters are \cite{}\todo{Cite matrix cookboox}:
\begin{equation}
\nabla_{\bm{\mu}} KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \bm{\Sigma}_p^{-1} (\bm{\mu} - \bm{\mu}_p)
\end{equation}
\begin{equation}
\nabla_{\bm{L}} KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \bm{\Sigma}_p^{-1}\bm{L} - (\bm{L}^{-1})^T
\end{equation}
Finally, the gradients w.r.t. the expected likelihood term of the variational objective \eqref{eq:elbo} can be computed using the reparameterization trick (e.g., \cite{}\todo{Cite deepmind and another}):
\begin{equation}
\nabla_{\bm{\mu}} \mathbb{E}_{\bm{w} \sim \mathcal{N}(\bm{\mu},\bm{LL}^T)}\left[ ||B(\bm{w})||_{D}^2 \right] = \mathbb{E}_{\bm{v} \sim \mathcal{N}(\bm{0},\bm{I})}\left[ \nabla_{\bm{w}} ||B(\bm{w})||_{D}^2 \right]\ \ \text{for}\ \ \bm{w} = \bm{Lv} + \bm{\mu}
\end{equation}
\begin{equation}
\nabla_{\bm{L}} \mathbb{E}_{\bm{w} \sim \mathcal{N}(\bm{\mu},\bm{LL}^T)}\left[ ||B(\bm{w})||_{D}^2 \right] = \mathbb{E}_{\bm{v} \sim \mathcal{N}(\bm{0},\bm{I})}\left[ \nabla_{\bm{w}} ||B(\bm{w})||_{D}^2 \cdot \bm{v}^T \right]\ \ \text{for}\ \ \bm{w} = \bm{Lv} + \bm{\mu}
\end{equation}

>>>>

Recent advances have allowed reinforcement learning (RL) \cite{sutton1998reinforcement} to achieve impressive results in a wide variety of complex tasks, ranging from Atari \cite{mnih2015human}, to the game of Go \cite{silver2016mastering}, to the control of sophisticated robotics systems \cite{kober2009policy,lillicrap2015continuous,levine2016end}. The main limitation is that RL algorithms still require an enormous amount of experience samples before successfully learning such complicated tasks. One of the most promising solutions is transfer learning, which focuses on reusing past knowledge available to the agent in order to reduce the sample-complexity for learning new tasks. In the typical settings of transfer in RL \cite{taylor2009transfer}, the agent is assumed to have already solved a set of \textit{source tasks} generated from some unknown distribution. Then, given a \textit{target task} drawn from the same distribution, or a slightly different one, the agent can rely on knowledge from the source tasks to speed-up the learning process. This constitutes a significant advantage over plain RL, where the agent learns each new task from scratch independently of previous learning experience. Several algorithms have been proposed in the literature to transfer experience samples \cite{lazaric2008transfer,taylor2008transferring}, policies/options \cite{fernandez2006probabilistic,konidaris2007building}, rewards \cite{konidaris2006autonomous}, features \cite{barreto2017successor}, parameters \cite{doshi2016hidden,killian2017robust}, and so on. We refer the reader to \cite{taylor2009transfer} for a thorough survey on transfer in RL.

@@

Under the assumption that tasks follow a certain distribution, an intuitive choice for designing a transfer algorithm is to attempt at characterizing the uncertainty over the target task. Then, an ideal algorithm would leverage prior knowledge from the source tasks to interact with the target task in such a way that this uncertainty is reduced as quickly as possible. This simple intuition makes Bayesian \todo{probabilistic? distributional?} methods appealing approaches for transfer in RL, and many previous works have been proposed in this direction. \cite{wilson2007multi} assume tasks share similarities in their dynamics and rewards and propose a hierarchical Bayesian model for the distribution of these two elements. Similarly, \cite{lazaric2010bayesian} assume tasks are similar in their value functions and design a different hierarchical Bayesian model for transferring such information. More recently, \cite{doshi2016hidden}, and its extension \cite{killian2017robust}, consider tasks whose dynamics are governed by some hidden parameters, and propose efficient Bayesian models for quickly learning such parameters in new tasks. However, all these algorithms require specific, and sometimes restrictive, assumptions (e.g., on the distributions involved or the function approximators adopted), which might limit their practical applicability. The importance of having transfer algorithms that alleviate the need of strong assumptions and can be easily adapted to different contexts motivates us to take a more general approach.

Similarly to \cite{lazaric2010bayesian}, we assume tasks to share similarities in their value functions and use the given source tasks to learn a distribution over such functions. Then, we use this distribution as a prior for learning the target task and we propose and efficient variational approximation of the corresponding posterior. Leveraging on recent ideas from randomized value functions \cite{osband2014generalization}, we design a Thompson sampling-based algorithm which efficiently explores the target task by repeatedly sampling from the posterior and acting greedily w.r.t. (with respect to) the sampled value function. We show that our approach is very general, in the sense that it does not require any specific choice of function approximator or prior/posterior distribution models. We propose a finite-sample analysis that theoretically validates our approach, while empirically evaluating its performance on three domains with increasing level of difficulty.

@@

We consider a distribution $\mathcal{D}$ over tasks, where each task is modeled as a discounted Markov decision process (MDP). We define an MDP as a tuple $\mathcal{M} = \langle \mathcal{S},\mathcal{A},\mathcal{P},\mathcal{R},p_0,\gamma\rangle$\todo{Should we provide a more transfer-oriented definition?}, where $\mathcal{S}$ is the state-space, $\mathcal{A}$ is a finite set of actions, $\mathcal{P}(\cdot | s,a)$ is the distribution of the next state $s'$ given that action $a$ is taken in state $s$, $\mathcal{R}: \mathcal{S}\times\mathcal{A}\rightarrow\mathbb{R}$ is the reward function, $p_0$ is the initial-state distribution, and $\gamma\in [0,1)$ is the discount factor. We assume the reward function to be uniformly bounded by a constant $R_{max}>0$. A deterministic policy $\pi : \mathcal{S} \rightarrow \mathcal{A}$ is a mapping from states to actions. At the beginning of each episode of interaction, the initial state $s_0$ is drawn from $p_0$. Then, the agent takes the action $a_0 = \pi(s_0)$, receives a reward $\mathcal{R}(s_0,a_0)$, transitions to the next state $s_1 \sim \mathcal{P}(\cdot | s_0,a_0)$, and the process is repeated. The goal is to find the policy maximizing the long-term return over a possibly infinite horizon: $\max_{\pi}J(\pi)\triangleq\mathbb{E}[\sum_{t=0}^{\infty} \gamma^t r_t \mid \mathcal{M},\pi]$. To this end, we define the optimal value function $Q^*(s,a)$ as the expected return obtained by taking action $a$ in state $s$ and following an optimal policy thereafter. Then, an optimal policy $\pi^*$ is a policy that is greedy with respect to the optimal value function, i.e., $\pi^*(s) = \argmax_a Q^*(s,a)$ for all states $s$. It can be shown (e.g., \cite{puterman1994markov}) that $Q^*$ is the unique fixed-point of the optimal Bellman operator $T$ defined by $TQ(s,a) = \mathcal{R}(s,a) + \gamma\mathbb{E}_{\mathcal{P}}[\max_{a'}Q(s',a')]$ for any value function $Q$. From now on, we adopt the term $Q$-function to denote any plausible value function, i.e., any function $Q : \mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ uniformly bounded by $\frac{R_{max}}{1-\gamma}$.

@@

Our approach is mostly related to \cite{lazaric2010bayesian}. Although we both assume the tasks to share similarities in their value functions, \cite{lazaric2010bayesian} consider only linear approximators and adopt a hierarchical Bayesian model of the corresponding weights' distribution, which is assumed Gaussian. On the other hand, our variational approximation allows for more general distribution families and can be combined with non-linear approximators. Furthermore, \cite{lazaric2010bayesian} propose a Dirichlet process model for the case where weights cluster into different classes, which relates to our mixture formulation and proves again the importance of capturing more complicated task distributions. In \cite{wilson2007multi}, the authors propose a hierarchical Bayesian model for the distribution over tasks. Differently from our approach and \cite{lazaric2010bayesian}, they consider a distribution over transition probabilities and rewards, rather than value functions. In the same spirit of our method, they consider a Thompson sampling-based procedure which, at each iteration, samples a new task from the posterior and solves it. However, \cite{wilson2007multi} consider only finite MDPs, which poses a severe limitation on the algorithm's applicability. On the contrary, our approach can handle high-dimensional tasks. In \cite{doshi2016hidden}, the authors consider a family of tasks whose dynamics are governed by some hidden parameters and use Gaussian processes (GPs) to model such dynamics across tasks. Recently, \cite{killian2017robust} extended this approach by replacing GPs with Bayesian neural networks, so as to obtain a more scalable approach.\todo{More comments about HiP-MDPs? Should we discuss more related works?}

In the RL community, our approach is related to value function randomization\cite{osband2014generalization}, which extends the well-known LSTD \cite{boyan1999least} by adopting Bayesian linear regression to model the uncertainty over the predicted value function weights, and use that to perform a form of Thompson sampling. Such approach was recently extended by \cite{azizzadenesheli2018efficient}, where the approximator is replaced by a Bayesian neural network, leading to an algorithm capable of solving much more complicated problems. Both these algorithms rely on the Gaussian assumption and, since they work in plain RL settings, have no informative prior available. On the other hand, our variational approximation allows more complex distributions (e.g., mixtures) to be adopted, while knowledge from the source tasks allows us to learn very informative priors.

@@

In this section, we provide an experimental evaluation of our approach in three different domains with increasing level of difficulty. In all experiments, we compare our Gaussian variational transfer algorithm (GVT) and the version using a $c$-component mixture of Gaussians ($c$-MGVT) to plain no-transfer RL (NT) with $\epsilon$-greedy exploration. To the best of our knowledge, no existing transfer algorithm is directly comparable to our approach from an experimental perspective. \todo{Any better motivation?}

@@

\begin{figure*}[t] 
  \begin{subfigure}[b]{0.15\textwidth}
    \includegraphics[trim=1.0cm 0.9cm 1.0cm 0.9cm,clip=true,height=2.0cm]{images/rooms.pdf}
    \caption{}
  \end{subfigure}
  \begin{subfigure}[b]{0.27\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.8cm 1.3cm,clip=true,height=2.9cm]{images/three-room/lrev.pdf}
    \caption{}
  \end{subfigure}
    \begin{subfigure}[b]{0.27\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.6cm 1.3cm,clip=true,height=2.9cm]{images/three-room-gen/lrev.pdf}
    \caption{}
  \end{subfigure}
    \begin{subfigure}[b]{0.27\textwidth}
    \includegraphics[trim=0.5cm 0cm 1.6cm 1.3cm,clip=true,height=2.9cm]{images/likelihood/lrew.pdf}
    \caption{}
  \end{subfigure}
  \caption{(a) the rooms environment, (b) transfer from $10$ source tasks with both doors moving, (c) transfer from $10$ source tasks with only one door moving, and (d) transfer performance as a function of how likely the target task is according to the prior.}
  \label{fig:dipfdbsgraphplots}
\end{figure*}

@@

In order to prove that our guesses are correct, we generate a set of $50$ source tasks for the three-room environment of Fig. \ref{} by sampling both door positions uniformly, and solve all of them by directly minimizing the TD error as presented in Sec. \ref{sec:td}. In order to make sure that their solutions are accurate enough, we allow sampling the initial state uniformly in the environment and run until converge. Then, we use our algorithms to transfer from $10$ source tasks sampled from the previously generated set. The average return over the last $50$ learning episodes as a function of the number of iterations is shown in Fig. \ref{}. Each curve is the result of $20$ independent runs, each resampling the target and source tasks. $95\%$ confidence intervals are shown. Further details on the parameters adopted in this experiment are given in App. \ref{}. As expected, the no-transfer algorithm NT fails at learning the task in so few iterations due to the limited exploration provided by an $\epsilon$-greedy policy. On the other hand, all our algorithms achieve a significant speed-up and are able to converge to the optimal performance in few iterations, with GVT being slightly slower. Interestingly, we notice that there is no advantage in adopting more than $1$ component for the posterior in MGVT. This is intuitive since, when the algorithm quickly figures out which task is being solved, it will move all components in the same direction.

To better understand the differences between GVT and MGVT, we now consider transferring from a slightly different distribution than the one from which target tasks are drawn. We generate again $50$ source tasks but this time with the bottom door fixed in the center and the other one moving. Then, we repeat the previous experiment, allowing both doors to move when sampling target tasks. The results are shown in Fig. \ref{}. Interestingly, MGVT seems almost unaffected by this change, proving that is has sufficient representation power to generalize to slightly different task distributions. The same does not hold for GVT, which now is not able to solve many of the sampled target tasks (this is the reason for the very high variance). This proves again that assuming Gaussian distributions can pose severe limitations in our transfer settings.

Finally, we analyze the transfer performance as a function of how likely the target task is according to the prior. We consider a two-room version of the environment of Fig. \ref{}. Differently from before, we generate tasks by sampling the door position from a Gaussian with mean $5$, and standard deviation $1.8$, so that tasks where the door is close to the sides are very unlikely. Fig. \ref{} shows the performance reached by GVT and MGVT with $1$ component at fixed iterations as a function of how likely the target task is according to such distribution. As expected GVT achieves poor performance on very unlikely tasks, even after many iterations. In fact, estimating a single Gaussian distribution definitely implies some information loss, especially about the unlikely tasks. On the other hand, MGVT keeps such information and, consequently, performs much better. Perhaps not surprisingly, MGVT reaches the optimal performance in $4k$ iterations no matter what task is being solved.

@@

@@

For this setting we propose a maze environment, such as those in Fig. *, for a robotic agent to navigate. We created a set of 20 mazes of size $10\times10$ with continuous space within which we ensured four groups: each group with the goal position in one of the corners of the maze. In order to simulate a robotic agent, we set an action space that allows to control the linear and angular motion by a fixed distance/angle each time an action is performed: \emph{move forward, rotate counter- and clockwise}. Moreover, we enhanced the state space to include, in addition to the absolute position and orientation, a set of distance measures to the nearest obstacles from the agent's perspective with maximum range of 2 and a field of vision from -90º to 90º relative to its orientation. In this way, we provide the agent with information that a robot would normally collect from LIDARs, sonars and other sensors alike. Finally, in the same field of perception, we provide a boolean vector indicating whether the goal is within the range of observation. Regarding rewards, the agent gets 1 whenever it enters the goal area and 0 in any other situation.

The main motivation for this set-up is that a robot is able to exploit locally-sensed information in addition to its estimation of the actual position---as opposed to the Rooms problem presented above---and it is such information that the robotic agent could learn to use and, also, exploit  in different maze configurations leveraging its previous experience. Moreover, simple variations in the maze structure could, potentially, produce big differences in the value function with respect to the ones seen in previous tasks which makes the setting very good to assess how robust our algorithm is robust to negative transfer.

These mazes were solved using Deep Q Networks (DQN) [] \todo{cite DQN?} with a multilayer perceptron with hidden layers $32\times32$  and, in order to perform the knowledge transfer, the target maze would be excluded from the source tasks given to the algorithm. Finally, the MGVT algorithm, with one component for the posterior distribution, was used to solve the target tasks shown in Fig. *.

@@

Under Gaussian distributions, all quantities of interest for using Alg. \ref{alg:transfer} can be computed very easily. The KL divergence between the prior and approximate posterior can be computed in closed-form as:
\begin{equation}
KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \frac{1}{2}\left( \log\frac{\left\lvert \bm{\Sigma}_p \right\rvert}{\left\lvert \bm{\Sigma} \right\rvert} + \text{Tr}\left( \bm{\Sigma}_p^{-1}\bm{\Sigma} \right) + (\bm{\mu} - \bm{\mu}_p)^T \bm{\Sigma}_p^{-1} (\bm{\mu} - \bm{\mu}_p) -K \right)
\end{equation}
for $\bm{\xi} = (\bm{\mu},\bm{L})$ and $\bm{\Sigma} = \bm{LL}^T$. Its gradients with respect to the variational parameters are:
\begin{equation}
\nabla_{\bm{\mu}} KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \bm{\Sigma}_p^{-1} (\bm{\mu} - \bm{\mu}_p)
\end{equation}
\begin{equation}
\nabla_{\bm{L}} KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \bm{\Sigma}_p^{-1}\bm{L} - (\bm{L}^{-1})^T
\end{equation}
Finally, the gradients w.r.t. the expected likelihood term of the variational objective \eqref{eq:elbo} can be computed using the reparameterization trick (e.g., \cite{hoffman2013stochastic,rezende2014stochastic}):
\begin{equation}
\nabla_{\bm{\mu}} \mathbb{E}_{\bm{w} \sim \mathcal{N}(\bm{\mu},\bm{LL}^T)}\left[ ||B(\bm{w})||_{D}^2 \right] = \mathbb{E}_{\bm{v} \sim \mathcal{N}(\bm{0},\bm{I})}\left[ \nabla_{\bm{w}} ||B(\bm{w})||_{D}^2 \right]\ \ \text{for}\ \ \bm{w} = \bm{Lv} + \bm{\mu}
\end{equation}
\begin{equation}
\nabla_{\bm{L}} \mathbb{E}_{\bm{w} \sim \mathcal{N}(\bm{\mu},\bm{LL}^T)}\left[ ||B(\bm{w})||_{D}^2 \right] = \mathbb{E}_{\bm{v} \sim \mathcal{N}(\bm{0},\bm{I})}\left[ \nabla_{\bm{w}} ||B(\bm{w})||_{D}^2 \cdot \bm{v}^T \right]\ \ \text{for}\ \ \bm{w} = \bm{Lv} + \bm{\mu}
\end{equation}
